working title

### 2024-06-17 writing/paper.qmd
writing progress

Prompt engineering defines the task of finding the right prompt to generate a desired output. Typically, this is an interactive and iterative process between human and LLM. There are certain writing techniques that improve the quality of
the LLM's response, such as providing examples for what the desired output should look like. There are various [blog posts](https://huggingface.co/docs/transformers/en/tasks/prompting#best-practices-of-llm-prompting) and books [e.g. @phoenix2024prompt] about how to write good prompts, in fact businesses are [hiring "Prompt Engineers"](https://indeed.com/q-prompt-engineer-jobs.html?vjk=98ac28acfd1328d7) even on dedicated [job boards](https://prompt-engineering-jobs.com/jobs/prompt-engineer-qwvmju/). Despite these, supposedly low-barrier, resources, end users without AI knowledge still struggle with prompt engineering [@zamfirescu2023johnny]. Not only can prompt engineering pose a [security risk](https://www.promptingguide.ai/risks/adversarial), but the process is also fuzzy, hard to replicate, and might still lead to unpredictable behaviors of the model. These issues pose risks for organizations, as the obscurity of manual prompt engineering can lead to loss of know-how, and the unpredictiability of the model can lead to lawsuits, damages to brands, or missed opportunities.

Problems of reproduceability, transparency, and documentation, become critical for organizations with the advent of the [EU AI Act](https://digital-strategy.ec.europa.eu/en/policies/regulatory-framework-ai). The act requires for "High risk applications", that the AI model features, e.g. "adequate mitigation systems", "traceability of results", and, "robustness". High risk applications are not common and important for society, they include applications e.g. in education, employment, and public services.

We propose a new type of numeric document summary, which captures all the information contained in the document, as it can be used to recreate the document itself. This document summary comes in the form of a numeric vector, which we call the summary embedding. To obtain this summary embedding, we maximize the likelihood of generating the document with a Large Language Model (LLM). In this sense, these summary embeddings are optimal. We show, that these embeddings capture inherent information about the document, can be used for classification, and generation of new documents. There are only practical requirements for estimating these summary embeddings, namely that backpropagation of gradients is possible from the likelihood to the input of the LLM.

Beyond the methodological contribution, we show a marketing application of these summary embeddings and make a step towards using Generative AI in market research and design of the marketing-mix. We find that we can use these embeddings to capture design motivations of marketers behind advertising claims, and establish a measure for linguistic uniqueness of an advertising claim. Suprisingly, we find that this linguistic uniqueness is negatively correlated with the perceived uniqueness of an advertising claim by consumers. We explore these linguistic features and find that claims with certain words and their synonyms, cluster together, e.g. claims that emphasize the taste of a yoghurt drink or present it as a breakfast. Furthermore, we show that claims that live in certain areas of a low dimensional representation of the summary embeddings, are more likely to be rated favorably by consumers, but that small differences in wording can lead to relatively large differences in evaluation. This suggests a two-step approach to designing a market research study for advertising claims, where the first stage should focus on a wide exploration of the design space, while the second stage should focus on a fine-grained evaluation of the most promising area in this design space. We end with a sketch on how these summary embeddings can be used in an AI augmented design process, similar to applications in image data by @burnap2023product and @ludwigMachineLearningTool2023.

- Consumer preferences in latent generation space & Ad-claims
  - Market research is expensive. How can we leverage these insights better to improve its benefits?
  - What is missing from preference learning? How is this different?
    - Media products/things are different from products as attribute bundles
      - We can directly represent the product itself?
    - Brands and their role in advertising
  - We can view products as attribute bundles and get utilities for these attributes 
  - How do we do this for products that cannot be "discretized"?
  - Generative AI for personalization
    - Previously: Set pieces, e.g. recommendation systems
    - True personalization: Generate solution for each individual consumer
  - TODO: Add more product claim specific
  - Advertisements can be effective, @daiFrontiersWhichFirms2023a observe an increase in purchase intentions for restaurants that adapt online advertising by 7%-19%.
  - they are relevant @kongMadeUSAClaims2021
  - costs
  - need flexibility in online settings
  - closely related problems, such as email targeting @sahniPersonalizationEmailMarketing2018
  - advertising is one of the marketing-mix instruments which companies use to position themselves in the market and shapes the incentives of the system itself @moorthyTargetingAdvertisingSpending2023 @renAdvertisingContentCreation2023
  - design question
  - Ban of cookie-based advertising due to GDPR, adapting to observable behavior in live sessions or to context on the website itself becomes more important. GDPR
  - Challenges to cookies in website morphing
  - @beckerDoesItPay2019, point to the nuanced role that authenticity plays in advertising, especially with respect to its alignment with the brand and product. 
  - @edelmanBrandingDigitalAge2010 Concept of the consumer decision journey, product claims could vary across multiple visits or encounters of the same customer

@@

@@

The BERT model is a transformer [see @vaswani2017attention] based model to represent text. Its developers have pre-trained BERT on a large corpus of text, using two different self-supervised learning objectives. The first objective is the so called "Masked Language Model" task. Here, the researchers replace a random selection of tokens in the text by a mask token, and the goal of the BERT model is to predict which token is underlying this mask. In order to learn about the information captures in a sentence, the researchers also pose the "Next Sentence Prediction" task to the model. For this training objective, the BERT model receives a pair of two sentences and needs to predict whether the second sentence follows the first one in a text. This task trains the `[CLS]` token: As it is pre-pended before every such sentence pair, it learns a representation of the information contained in the focal sentence. In practice, the output for a `[CLS]` token serves as a powerful feature e.g. for sentence classification. BERT embeddings are versatile, but can also be fine-tuned to a specific task such as text summarization. However, such a fine-tuning requires the user to compile a dataset of text-summary and text pairs, that are representative of their application. These BERT representations are only descriptive and cannot be used for the generation of new text [see @devlinBERTPretrainingDeep2018]. There exist expansions to BERT, such as RoBERTa by @yinhanliuRoBERTaRobustlyOptimized2019, which improve the optimization procedure.

@@

We can also summarize documents by passing these to an LLM and prompting the model to write a summary for us [e.g. @huangenhanced; @chakrabortyAbstractiveSummarizationEvaluation2024]. However, these summaries are textual, not deterministic, and their quality depends on the formulation of the prompt. For example, @liuLostMiddleHow2023 find that answers of LLMs are better, when we place relevant information at the beginning or end of the prompt, while inserting emotions into prompts also improves the quality of the LLMs response, in some cases even doubling its performance [see @liLargeLanguageModels2023]. Another issue is that these summaries can vary in length, and one needs to specify how detailed a summary should be. The difference in length between the summary and focal document, also makes it challenging to evaluate how much of the information of the focal document is captured by the summary [@chakrabortyAbstractiveSummarizationEvaluation2024]. If these summaries are of a high quality, i.e. represent the information in the focal document well, then they can be useful when we want to share a written summary of a document. However, for many data science and automation tasks, we need a reliable numeric representation of the document, which is the focus of this paper.

@@

| Method                              | Origin                                          | Reference                              |  Deterministic? |  Generation?  |  Type?  |
|:------------------------------------|:------------------------------------|:--------------------|:----------------------:|:----------------------:|:----------------------:|
| BERT `[CLS]` token                       | Next-sentence prediction task                  | @devlinBERTPretrainingDeep2018         | $\checkmark$   | $\times$       | Numeric    |
| Pooled Word Embeddings              | Aggregation of token information               | @shenBaselineNeedsMore2018             | $\checkmark$   | $\times$       | Numeric    |
| Prompt engineering                  | Emergent capability of LLM                     | CITE                                   | $\times$       | $\times$       | Textual    |
| Reverse engineered document summaries | Maximize the likelihood to re-generate the focal document | *Proposed method*                        | $\checkmark$   | $\checkmark$   | Numeric |
: Overview of existing methods to summarize documents. {#tbl-docsum}

@@

<!-- TODO: Normalization layer and its role; Does it work without it? Why / not? -->

Generative Large Language Models (LLM), such as [ChatGPT](https://chat.openai.com/) [@radford2018improving], generate text based on textual inputs, so called prompts. Internally, the model translates the prompt first into tokens, integer codes representing certain words, symbols and syllabuls, and then into a lower dimensional vector representation of the text itself, which we call the input embedding. This input embedding is then passed through a neural network that yields a probability distribution over the next token in the sequence. To generate text, the model makes a draw from this distribution, and attached this generated token at the end of the prompt, repeating this process until it predicts the next token to be the end-of-text token (eos), i.e. until it predicts that the text ends. LLMs are pre-trained on large text corpora, such as the [Common Crawl](https://commoncrawl.org/) which contains over 250 billion pages of text. Their architecture is rooted in the attention mechanism. The attention mechanism models interactions in the text, such as negations, synonyms, or grammatical structures, across a long sequence of text. @vaswani2017attention implement this mechanism in the transformer block, which is a modular building block for the neural network underlying an LLM. The combination of these transformer blocks, giant pre-training data, and huge computing resources, lead to LLMs that have "emergent capabilities". These are capabilities, which were not explicitly trained for by the developers of the model, such as the ability of a model to translate between languages, pass the BAR exam, or write textual summaries of documents [see @weiEmergentAbilitiesLarge; @luAreEmergentAbilities2023]. Unidirectional LLMs, such as GPT or LLaMa [@radford2018improving; @metaMetaLlama], are built for text-generation. They predict the next token in a sequence of tokens autoregressively, therby, these models only work from left-to-right (or right-to-left). Bidirectional LLMs, such as BERT [@devlinBERTPretrainingDeep2018], are built for representing text. They predict a token given the context around this token, and can hence read the text both from left-to-right and right-to-left. Both types of models share two note-worthy components, that account for a large share of their power. One, developers pre-train both types of models on large bodies of text, to learn about the structure of language itself. Two, by using a transformer architecture [@vaswani2017attention], which enables these language models to learn dependencies between words, such as negations (short-range dependency), or document structures, e.g. salutations in letters (long-range dependency). <!-- TODO: Simplify this paragraph -->

- For generation of sequences, the LLM autoregressively predicts the next token, given the previous tokens. At each step in the generation, a selection of the next token needs to take place, for which there exist various procedures (CITE). As we are optimizing with respect to the joint likelihood of the target sequence, we also need to generate with a procedure that generates this most likely sequence. Finding this sequence from the generation tree is a combinatorical problem and computationally not feasible. Hence, we use the beam-search heuristic to make our generation. Rather than only tracking the most likely token at each generation step, beam search also tracks the $n$-most likely tokens. The more beams we search, the more likely we are to find a sequence with a higher likelihood, then we would have found otherwise [@huggingfaceGenerateText].

@@

As we are interested in finding a low-dimensional sub-space, in which we can represent the $D$ advertising claims, we use a factor model and estimate the summary embeddings for all advertising claims jointly, yielding the $D \times E$ matrix of summary embeddings $\mathbf{S}$. We restrict the degrees of freedom to $F$ hidden factors. To implement this factor model, we create an Autoencoder with one hidden layer, that is fully connected with linear activation functions [@goodfellowDeepLearning2016]. The input data to this autoencoder is an identity matrix of dimension $D$. We illustrate this neural network representation of the factor model in @fig-factor.

@@

We denote the matrix of weights of the encoder and decoder layers by $\mathbf{W}^{Encoder}$ and $\mathbf{W}^{Decoder}$. These matrices have the dimensions $D \times F$ and $F \times E$ respectively. We also define the vectors of biases for the encoder and decoder by $\boldsymbol{b}^{Encoder}$ (length $F$) and $\boldsymbol{b}^{Decoder}$ (length $E$), and stack them into matrices by taking the Kronecker product with a length $D$ vector of ones, $\boldsymbol{1}_D$, to obtain the biases in matrix form as $\mathbf{B}^{Encoder} = \boldsymbol{1}_D \otimes \boldsymbol{b}^{Encoder}$ and $\mathbf{B}^{Decoder} = \boldsymbol{1}_F \otimes \boldsymbol{b}^{Decoder}$. For notational convenience, we denote this factor model by $\operatorname{Factor}_{\Omega_{Factor}} \left(\mathbb{I}_{D} \right): \lbrace 0, 1 \rbrace^{D \times D} \to \mathbb{R}^E$, with parameters $\Omega_{Factor}$, where $\Omega_{Factor} = \lbrace \mathbf{W}^{Encoder}, \mathbf{W}^{Decoder}, \mathbf{B}^{Encoder}, \mathbf{B}^{Decoder} \rbrace$.

@@

<!-- TODO: Can I split this into separate optimizations? Why / why not? -->
$$
 \mathbf{S}^{*} = \argmax_{\mathbf{W}^{Encoder}, \; \mathbf{W}^{Decoder}, \; \mathbf{B}^{Encoder}, \; \mathbf{B}^{Decoder}} \operatorname{\mathcal{L}}_D \left ( \mathbf{S} \right) \text{. }
$$

@@

<!-- TODO: What does it really return? -->
<!-- TODO: Good idea or not? -->
<!-- TODO: Figure of generation tree -->

@@

CITE: Prompt engineering

@@

As as robustness check, we repeated this task for different attributes (instead of shiny, healthiness and colorfulness of hair), and for a different product category (surface cleaners). In total, we obtain 80 advertising claims, 20 for each of these settings. Half of these claims are supposed to be tangible and half intangible.

We obtain data from a market research company, that includes advertising claims for different FMCG. In our application, we focus on dairy and hair shampoo products. These data, stem from different variations of choice-based conjoint studies [@eggersChoiceBasedConjointAnalysis2022], which we cannot further disclose due to confidentiality agreements. The data includes choice-experiments aggregated at the level of the advertising claims, hence we do not have responses of individuals. These measurements of consumer-preferences are on multiple dimensions, such as relevance, uniqueness of a claim, brand fit, or an overarching rating of the claim. The advertising claims were designed for markets in different countries and are in english (US, UK) or translated into english (others). The date of the study is also included in the data. We also have information on the marketers motivation behind the design of an advertising claims, such as whether she designed the claims to pronounce the health benefits of a yoghurt, or the taste yoghurt, as well as the brand of the product.

@@

- Descriptives of the data

- Prepocessing of the data

@@

@@

TODO: Have similarity measure vs euclidean distance in encoding space

[ ] Training of embeddings

- computation
- dealing with spikes

[ ] Analysis of embeddings; Plot correlation of claims and dims, PCA

::: {#fig-correlation layout-ncol=2 layout-nrow=2}
![RE](figures/correlation_re.pdf)

![BoW](figures/correlation_bow.pdf)

@@

::: {#fig-correlation-surface layout-ncol=2 layout-nrow=2}
![RE](figures/correlation_re_surface.pdf)

![BoW](figures/correlation_bow_surface.pdf)

>>>>

Prompt engineering defines the task of finding the right prompt to generate a desired output. Typically, this is an interactive and iterative process between human and LLM. There are certain writing techniques that improve the quality of
the LLM's response, such as providing examples for what the desired output should look like. There are various [blog posts](https://huggingface.co/docs/transformers/en/tasks/prompting#best-practices-of-llm-prompting) and books [e.g. @phoenix2024prompt] about how to write good prompts, in fact businesses are [hiring "Prompt Engineers"](https://indeed.com/q-prompt-engineer-jobs.html?vjk=98ac28acfd1328d7), even on dedicated [job boards](https://prompt-engineering-jobs.com/jobs/prompt-engineer-qwvmju/). Despite these, supposedly low-barrier, resources, end users without AI knowledge still struggle with prompt engineering [@zamfirescu2023johnny]. Not only can prompt engineering pose a [security risk](https://www.promptingguide.ai/risks/adversarial), but the process is also fuzzy, hard to replicate, and might still lead to unpredictable behaviors of the model. These issues pose risks for organizations, as the obscurity of manual prompt engineering can lead to loss of know-how, and the unpredictiability of the model can lead to lawsuits, damages to brands, or missed opportunities. Problems of reproduceability, transparency, and documentation, become critical for organizations with the advent of the [EU AI Act](https://digital-strategy.ec.europa.eu/en/policies/regulatory-framework-ai). The act requires for "High risk applications", that the AI model features, e.g. "adequate mitigation systems", "traceability of results", and, "robustness". High risk applications are not common and important for society, they include applications e.g. in education, employment, and public services.

We propose a new type of numeric document summary, which captures all the information contained in the document, as it can be used to recreate the document itself. This document summary comes in the form of a numeric vector, which we call the summary embedding. To obtain this summary embedding, we maximize the likelihood of generating the document with a Large Language Model (LLM). In this sense, these summary embeddings are optimal. To mitigate overfitting and to make these embeddings more interpretable, we also introduce a factor model to estimate these document summaries. We show, that these embeddings capture inherent information about the document, can be used for classification, and generation of new documents. There are only practical requirements for estimating these summary embeddings, namely that backpropagation of gradients is possible from the likelihood to the input of the LLM.

Beyond the methodological contribution, we show a marketing application of these summary embeddings by applying them to market research data on product claims for yoghurt drinks. We find that we can use these embeddings to capture design elements of advertising claims, and establish a measure for linguistic uniqueness of an advertising claim. Suprisingly, we find that this linguistic uniqueness is negatively correlated with the perceived uniqueness of an advertising claim by consumers. We explore these linguistic features and find that claims with certain words and their synonyms, cluster together, e.g. claims that emphasize the taste of a yoghurt drink or present it as a breakfast. Furthermore, we show that claims that live in certain areas of a low dimensional representation of the summary embeddings, are more likely to be rated favorably by consumers, but that small differences in wording can lead to relatively large differences in evaluation. This suggests a two-step approach to designing a market research study for advertising claims, where the first stage should focus on a wide exploration of the design space, while the second stage should focus on a fine-grained evaluation of the most promising area in this design space. We end with a sketch on how these summary embeddings can be used in an AI augmented design process, similar to applications in image data by @burnap2023product and @ludwigMachineLearningTool2023.

We proceed with a brief overview of the relevant literature and introduce our methodology. After testing our procedure on synthetic data, and exploring the low-dimensional factor space for the generation of new advertising claims, we apply our model to the market research data. We conclude with managerial implications, a discussion of the limitations of our approach and expansions for this research.

@@

<!-- TODO: Merge these paragraphs -->

LLMs predict the next word in a sequence of words. These models, such as [ChatGPT](https://chat.openai.com/) [@radford2018improving], generate text based on textual inputs, so called prompts. Internally, the model translates the prompt first into tokens, integer codes representing certain (sequences of) characters, and then into a lower dimensional vector representation of the text itself, which we call the input embedding. This input embedding is then passed through a neural network that yields a probability distribution over the next token in the sequence. To generate text, the model makes a draw from this distribution, and attached this generated token at the end of the prompt, repeating this process until it predicts the next token to be the end-of-text token (eos), i.e. until it predicts that the text ends. LLMs are pre-trained on large text corpora, such as the [Common Crawl](https://commoncrawl.org/) which contains over 250 billion pages of text. Their architecture is rooted in the attention mechanism. The attention mechanism models interactions in the text, such as negations, synonyms, or grammatical structures, across a long sequence of text. @vaswani2017attention implement this mechanism in the transformer block, which is a modular building block for the neural network underlying an LLM. The combination of these transformer blocks, giant pre-training data, and huge computing resources, lead to LLMs that have "emergent capabilities". These are capabilities, which were not explicitly trained for by the developers of the model, such as the ability of a model to translate between languages, pass the BAR exam, or write textual summaries of documents [see @weiEmergentAbilitiesLarge; @luAreEmergentAbilities2023].

@@

The BERT model is a transformer [see @vaswani2017attention] based model to represent text. Its developers have pre-trained BERT on a large corpus of text, using two different self-supervised learning objectives. The first objective is the so called "Masked Language Model" task. Here, the researchers hide a random token in a sequence, and the goal of the BERT model is to predict which token is missing. This teaches the model the meaning of words in context. In order to learn about the information captured in a sentence, the researchers also pose the "Next Sentence Prediction" task to the model. For this training objective, the BERT model receives a pair of two sentences and needs to predict whether the second sentence follows the first one in a text. This task trains the `[CLS]` token: As it is pre-pended before every such sentence pair, it learns a representation of the information contained in the focal sentence. In practice, the output for a `[CLS]` token serves as a powerful feature e.g. for sentence classification. BERT embeddings are versatile, but can also be fine-tuned to a specific task such as text summarization. Such a fine-tuning requires the user to compile a dataset of summary-text pairs, that are representative of their application. These BERT representations are only descriptive and cannot be used for the generation of new text [see @devlinBERTPretrainingDeep2018]. There exist improvements to BERT, such as RoBERTa by @yinhanliuRoBERTaRobustlyOptimized2019, which improve the optimization procedure.

@@

We can also summarize documents by passing these to an LLM and prompting the model to write a summary for us [e.g. @huangenhanced; @chakrabortyAbstractiveSummarizationEvaluation2024]. However, these summaries are textual, not deterministic, and their quality depends on the formulation of the prompt. For example, @liuLostMiddleHow2023 find that answers of LLMs are better, when we place relevant information at the beginning or end of the prompt, while inserting emotions into prompts also improves the quality of the LLMs response, in some cases even doubling its performance [see @liLargeLanguageModels2023]. Another issue is that these summaries can vary in length, and one needs to specify how detailed a summary should be. The difference in length between the summary and focal document, also makes it challenging to evaluate how much of the information of the focal document is captured by the summary [@chakrabortyAbstractiveSummarizationEvaluation2024]. If these summaries are of a high quality, i.e. represent the information in the focal document well, then they can be useful when we want to share a written summary of a document. However, for many data science and automation tasks, we need a reliable numeric representation of the document, which is the focus of this paper. We could use these textual summaries for generation, in the sense that we could prompt the model to generate a new document based on the summary. A recent contribution, that can make prompt engineering more reliable is the framework proposed by @khattabDSPyCompilingDeclarative2023, called DSPy, which automatically learns how to combine different prompting and finetuning techniques, in order to improve the generated answer of the LLM with respect to a pre-defined metric. They show that their discrete optimization procedure leads to better generated output compared to few-shot prompting and expert designed prompts, even when applied to smaller LLMs. Their approach is text based and revolves around generating examples of the desired output, and then tuning the prompt and the LLM itself based on these examples. In spirit, their approach is similar to hyper-parameter tuning, in that it uses a type of (small) training data and a performance metric (e.g. whether the answer is an exact match to a certain label), and then adjusts the prompt and the LLM to maximize the scoring of its answer on this training set. The twist is, that the DSPy framework can perform such an hyper-parameter tuning in an automized fashion, by generating candidate prompts itself and selecting from them. Our proposed approach differs from DSPy, in that we optimize in a continuous space, and that our goal is to find a numeric summary for a document, rather than to generate a good response to natural language tasks, such as question answering.

@@

| Method                              | Origin                                          | Reference                              |  Deterministic? |  Generation?  |  Type?  |
|:------------------------------------|:------------------------------------|:--------------------|:----------------------:|:----------------------:|:----------------------:|
| BERT `[CLS]` token                       | Next-sentence prediction task                  | @devlinBERTPretrainingDeep2018         | $\checkmark$   | $\times$       | Numeric    |
| Pooled Word Embeddings              | Aggregation of token information               | @tomasmikolovDistributedRepresentationsWords2013 @shenBaselineNeedsMore2018             | $\checkmark$   | $\times$       | Numeric    |
| (Optimized) Prompt engineering                  | Emergent capability of LLM                     | @khattabDSPyCompilingDeclarative2023; @huangenhanced                                   | $\times$       | $\checkmark$       | Textual    |
| Reverse engineered document summaries | Maximize the likelihood to re-generate the focal document | *Proposed method*                        | $\checkmark$   | $\checkmark$   | Numeric |
: Overview of existing methods to summarize documents. {#tbl-docsum}

@@

<!-- TODO: Normalization layer and its role; Does it work without it? Why / not? -->
<!--

- For generation of sequences, the LLM autoregressively predicts the next token, given the previous tokens. At each step in the generation, a selection of the next token needs to take place, for which there exist various procedures (CITE). As we are optimizing with respect to the joint likelihood of the target sequence, we also need to generate with a procedure that generates this most likely sequence. Finding this sequence from the generation tree is a combinatorical problem and computationally not feasible. Hence, we use the beam-search heuristic to make our generation. Rather than only tracking the most likely token at each generation step, beam search also tracks the $n$-most likely tokens. The more beams we search, the more likely we are to find a sequence with a higher likelihood, then we would have found otherwise [@huggingfaceGenerateText].
  -->

@@

As we are interested in finding a low-dimensional sub-space, in which we can represent the $D$ advertising claims, we use a factor model and estimate the summary embeddings for all advertising claims jointly, yielding the $D \times E$ matrix of summary embeddings $\mathbf{S}$. We restrict the degrees of freedom to $F$ hidden factors. To implement this factor model, we create an Autoencoder with one hidden layer, that is fully connected with linear activation functions [@goodfellowDeepLearning2016]. The input data to this autoencoder is an identity matrix of dimension $D$. Before passing the summary embedding to the LLM, we pass these nodes into a Layer Normalization to stabilize the training process [@baLayerNormalization2016]. Layer normalization normalizes across the hidden nodes, rather than across the observations in a batch [batch-normalization, see @ioffeBatchNormalizationAccelerating2015]. Thereby, it can also be used with a single observation. We illustrate this neural network representation of the factor model in @fig-factor.

@@

We denote the matrix of weights of the encoder and decoder layers by $\mathbf{W}^{Encoder}$ and $\mathbf{W}^{Decoder}$. These matrices have the dimensions $D \times F$ and $F \times E$ respectively. We also define the vectors of biases for the encoder and decoder by $\boldsymbol{b}^{Encoder}$ (length $F$) and $\boldsymbol{b}^{Decoder}$ (length $E$), and stack them into matrices by taking the Kronecker product with a length $D$ vector of ones, $\boldsymbol{1}_D$, to obtain the biases in matrix form as $\mathbf{B}^{Encoder} = \boldsymbol{1}_D \otimes \boldsymbol{b}^{Encoder}$ and $\mathbf{B}^{Decoder} = \boldsymbol{1}_F \otimes \boldsymbol{b}^{Decoder}$. For notational convenience, we denote this factor model by $\operatorname{Factor}_{\Omega_{Factor}} \left(\mathbb{I}_{D} \right): \lbrace 0, 1 \rbrace^{D \times D} \to \mathbb{R}^{D \times E}$, with parameters $\Omega_{Factor}$, where $\Omega_{Factor} = \lbrace \mathbf{W}^{Encoder}, \mathbf{W}^{Decoder}, \mathbf{B}^{Encoder}, \mathbf{B}^{Decoder} \rbrace$.

@@

<!-- TODO: Is it S or S_{d, :}? Can I split this into separate optimizations? Why / why not? -->
$$
 \mathbf{S}^{*} = \argmax_{\mathbf{W}^{Encoder}, \; \mathbf{W}^{Decoder}, \; \mathbf{B}^{Encoder}, \; \mathbf{B}^{Decoder}} \operatorname{\mathcal{L}}_D \left ( \mathbf{S} \right) \text{. }
$$

@@

<!-- TODO: Talk about Layer Normalization, explain difference to BatchNormalization -->
<!-- TODO: Why does result depend on Layer Normalization? Is there a bug, why not? (Scrambled version works too) -->

@baLayerNormalization2016

- also used in @vaswani2017attention
- can be used in test data
- would like not to use it
- majority of claims close to zero; However, some outliers which are pushed away from the others

<!-- TODO: Note on optimization and behavior of likelihood, spike in gradient -->
<!-- TODO: What does it really return? -->
<!-- TODO: Good idea or not? -->
<!-- TODO: Figure of generation tree -->

## Diagnostics

- Other methods as benchmarks
  - Word2Vec
  - BERT

To analyze our results, we will use a suite of diagnostic tools, which we will explain here briefly.

- Correlation matrix

  - First, we de-mean each embedding dimension by subtracting the mean of this embedding dimension calculated across observations. The motivation for this is, to take out parts of the embedding that are due to the domain of the advertising claims.

  - Correlation matrices across the claims and across the embedding dimensions: When analyzing the summary embeddings, we form correlations across the advertising claims and correlations across the embedding dimensions. The former is the same as a correlation matrix of a data matrix with as many rows as embedding dimensions and as many columns as the number of observations, while the latter is the correlation matrix of a data matrix with as many rows as observations and as many columns as embedding dimensions.

- Leave-one-out regression (LOO)
  - We also perform leave-one-out regression, here we regress a focal embedding onto the average embedding from the one class and the average embedding from the other class, and an intercept. We repeat this for all observations. Since all embeddings are de-meaned, we expect an average intercept of zero for all regressions. We also expect that the estimates for the focal claim's group are positive, while the estimates for the other group should be negative.

- Principal Component Analysis (PCA)
  - PCA is a dimensionality reduction technique, which is rooted in the singular value decomposition. Intuitively, it projects the data onto orthogonal axes in a way that each axis captures as much variance as possible. We select the first two principal components, as this allows us to visualize our data.
  - Eigenvalue? Scree plot?

@@

As a first evaluation, we create a small synthetic dataset of 20 advertising claims for a haircare product by querying ChatGPT with the prompt in @fig-gpt. We want these claims to differ whether they are pronouncing tangible or intangible attributes of the product. These advertising claims all focus on how "shiny" the product ones hair makes. As as robustness check, we repeated this task for different attributes (instead of shinyness: Healthiness and colorfulness of hair), and for a different product category (surface cleaners). In total, we obtain 80 advertising claims, 20 for each of these settings. Half of these claims are supposed to be tangible and half intangible. Additionally, we check whether our results are robust to changes in the order of the claims, to rule out data leakage. We present these data in @tbl-claims.

@@

For our empirical application, we obtain data from a market research company, that includes advertising claims for Yoghurt Drinks. These data, stem from different variations of choice-based conjoint studies [@eggersChoiceBasedConjointAnalysis2022], which we cannot further disclose due to confidentiality agreements. The data includes choice-experiments aggregated at the level of the advertising claims, hence we do not have responses of individuals. These measurements of consumer-preferences are on multiple dimensions, such as relevance, uniqueness of a claim, brand fit, or an overarching rating of the claim. The advertising claims were designed for markets in different countries and are in english (US, UK) or translated into english (others). The date of the study is also included in the data. All products are from the same brand. In total, we have 60 observations, 20 for each country. Many of the researched advertising claims occurr in multiple of these separate studies. In total, we have 32 unique advertising claims. Most advertising claims get a uniqueness measure around 5.8, with the whole distribution ranging from 4.8 to 7.4. When we look at the uniqueness measure with respect to the country in which the study was held, we see that the variance is the smallest for the German market and the largest for the spanish market. However, the means are not statistically different across the countries. The distribution of the rating measure is similar in shape, ranging from 43 to 80. The rating and uniqueness are positively correlated (correlation of $0.8843$)

<!-- We also have information on the marketers motivation behind the design of an advertising claims, such as whether she designed the claims to pronounce the health benefits of a yoghurt, or the taste yoghurt, as well as the brand of the product.

- One brand, yoghurt drink
- MaxDiff experiment, ranking of choices
- Relevance, fit with brand, rank
- 32 unique claims, 3 countries: Germany, Spain, UK

@@

The scale for the uniqueness measure is from 0 to 10 and for the rating measure from 0 to 100.

::: {#fig-eda-mr4-uni layout-ncol=2}
![Histogram across countries](figures/uniqueness_mr4_eda.pdf){#fig-eda-mr4-uni-a}

![Boxplot per country](figures/uniqueness_box_mr4_eda.pdf){#fig-eda-mr4-uni-b}

Distributions of measure for claim uniqueness and ratings across countries.
:::

::: {#fig-eda-mr4-rating layout-ncol=2}
![Histogram across countries](figures/rating_mr4_eda.pdf)

![Boxplot per country](figures/rating_box_mr4_eda.pdf)

Distributions of measure for claim uniqueness and ratings across countries.
:::

![Scatterplot of Uniqueness and Rating measures, colored by country.](figures/scatterplot_mr4_eda.pdf)

@@

<!-- ## MR 6

- 5 countries: Brazil, Mexico, Russia, France, Germany
- 40 unique claims -->

@@

## Single Embeddings

- HairColor: 42 minutes
- HairHealth: 14 minutes
- Surface: 16 minutes
- Surface Scramble: 9 minutes

<!-- TODO: fill in the computation times -->
<!-- TODO: Have model with e.g. 50 factors? Then for many claims? -->
<!-- TODO: Have model without LayerNorm -->

For this project, we obtained a [SURF NWO Small Compute Grant](https://www.surf.nl/en/small-compute-applications-nwo). We train these summary embeddings until we obtain a likelihood to generate the target claim of $0.99$. Performing this training on a Nvidia A100-GPU, takes about 30 seconds for all 20 claims. The training of the factor representation with 2 factors and for the same claims is more computationally intense and takes about Y minutes on the same computer. Here we traing for a joint likelihood of $0.99$. We always verified that the summary embeddings regenerate all focal claims correctly.

@fig-correlation-single

::: {#fig-correlation-single layout-ncol=3}
![RE](figures/correlation_re_hair_single.pdf)

@@

- does not work to capture the two classes: Overfitting?
- does regularization help?

We can interpolate between to target strings by forming a weighted average of their summary embeddings. @tbl-interpolation shows the strings that we generated for a convex combination of the claims *"Unlock the secret to hair that shines from within, reflecting your inner glow."* (A) and *"Rediscover the joy of hair that beams with inner vibrancy."* (B). We combine them as $\boldsymbol{s}_{combine} = weight * \boldsymbol{s}_{A} + (1 - weight) * \boldsymbol{s}_{B}$, where $weight \in [0, 1]$. For weights that are close to either $0$ or $1$ we still generate the respective focal claim. When we start to step further into the middle between the two claims, first the end of the sequence changes (e.g. *"Unlock the secret to blackened nails that shine from the shine."* at for $weight = \frac{3}{20}$). For weights that are close to $\frac{1}{2}$, we stop generating eos-tokens and the generated strings become unintelligble. However, these strings still maintain some of the structure of the focal claims, such as having words related to beauty products and starting with the syllable "Un" or "Red".

|    | Weight | Generated String                                                                                      |
|---:|:-------|:------------------------------------------------------------------------------------------------------|
|  0 | 0/20   | Unlock the secret to hair that shines from within, reflecting your inner glow.   \textit{eos-token}                     |
|  1 | 1/20   | Unlock the secret to hair that shining from within, your bedroom. \textit{eos-token}                                     |
|  2 | 2/20   | Unlock the secret to black metal's glow. \textit{multiple eos-token}                                                              |
|  3 | 3/20   | Unlock the secret to blackened nails that shine from the shine.   \textit{eos-token}                                    |
|  4 | 4/20   | Un to of for the team members upon                                                                    |
|    |        | Moderation of the, that                                                                               |
|  5 | 5/20   | Un to (93) of 15 + 18 + 36                                                                            |
|    |        | Add to                                                                                               |
|  6 | 6/20   | Un to (mit. of private eye and nirvana-ly-ly                                                          |
|  7 | 7/20   | Uncle-healedered with a lifetime-changing blend of light and energy and                               |
|  8 | 8/20   | Unclelyed by the Tone's Bright Side™ Lipstick.                                                        |
|    |        | Bright                                                                                                |
|  9 | 9/20   | Unclelying our clients and fory, we're sure about your smile.                                         |
| 10 | 10/20  | Redmates on the Road to of Dreams.                                                                    |
|    |        | Inspirating your soul with                                                                            |
| 11 | 11/20  | Redmates at the Sunlight Your Hair.     \textit{multiple eos-token}                                                              |
| 12 | 12/20  | Redhs founder, that a master the way she loves......                                                  |
| 13 | 13/20  | Redhs founder, of where, at, and to and to, and can be                                                |
| 14 | 14/20  | Rediscover the joy of purpose and detail with a strand of your hair. (and                             |
| 15 | 15/20  | Rediscover the joy of success with long,istry's health and creativity. back to                        |
| 16 | 16/20  | Rediscover the joy of hair that loves to making you.™ the way.br                                      |
| 17 | 17/20  | Rediscover the joy of hair that's at the-corrects life with a healthy                                 |
| 18 | 18/20  | Rediscover the joy of hair that beams with inner vibrancy. \textit{eos-token}                                           |
| 19 | 19/20  | Rediscover the joy of hair that beams with inner vibrancy. \textit{eos-token}                                           |
: Interpolation between two advertising claims. {#tbl-interpolation}

## Factor Representation
